\textbf{Replies to Reviewers' Comments}

### 2023-06-28 Articles/2022-ActionPrinciple/ScientificReportsAmended/response1.tex
Minor edits to reviewer responses before submitting them

\begin{quote}
 Some of the standard notation and terminology of differential geometry are not present in the paper. For example, the discussion in the supplementary document from line 115 to 124 could be referred to as the volume preserving diffeomorphism of flow maps that are presented in Kambe, Tsutomu Jixin. Geometrical Theory Of Dynamical Systems And Fluid Flows (Revised Edition). Vol. 23. World Scientific Publishing Company, 2009. Also, for standard notation of differential geometry, please see Kreyszig, Erwin. Differential geometry. Courier Corporation, 2013. 
\end{quote}
As we stated in the supplementary information of the original version, the use of standard notation is avoided on purpose because, in our experience, it does map well to physical intuition. We added an additional comment to that effect in the main body of the paper on line 26, as we realized this was not made sufficiently clear. In response to the other reviewer, we have also clarified in the paper the connection between the mathematical objects $\vec{S}$, $\omega$ and $\theta$ in our paper and in the standard mathematical literature, which further highlights the mismatch between standard mathematical definitions and the physically motivated ones. For example, $\theta$ would correspond to the canonical one form in the context of symplectic geometry and the contact form in the context of contact manifolds. Both of these objects are uniquely defined mathematically. However, given that the physical object we represent with $\theta$ is a vector potential, it is not uniquely defined as it depends on the arbitrary gauge choice. Therefore the ``mathematical'' $\theta$s represent the vector potential with a specific gauge choice.

@@

\begin{quote}
 General comments and questions
 Is the study limited to divergence free vector fields only? 
\end{quote}
Yes. The study is limited to all cases where the principle of stationary action holds (i.e. yields unique solutions) which, by virtue of the Liouville theorem, corresponds to divergence free displacement vector fields.

@@

\begin{quote}
 Why call the Principle of Least Action the Action Principle? 
\end{quote}
We use the term in the title to not make it exceedingly long. It is not an uncommon abbreviation. For example, the book ``The action principle and partial differential equations'' by D. Christodoulou.  We generally use ``principle of stationary action'' throughout the paper.

@@

 In pp. 11 line 140 in the supplementary document, “Putting …” should be “Summing or Grouping …”.
\end{quote}
We thank the reviewer for taking the time to point out the typos, which have been fixed. We also have reworded the sentence.

@@

\begin{quote}
	Now, I am curious, and would like the authors to clarify what mathematical objects are new.  The displacement field S was new to me. Is the potential $\theta$, also new?  And the counting form, lower case $\omega$, is this the symplectic 2-form familiar from Hamiltonian geometry?  If so, please state. And if not, perhaps a different symbol would be better.
\end{quote}
We added remarks when the symbols are introduced in the main text, at lines 41, 52 and 98. We also added more explanation in the supplementary material at lines 186, 204 and 380. The displacement field S is typically referred to as "the vector field" in dynamical system theory. $\theta$ roughly corresponds to the canonical one form/contact form and $\omega$ to the symplectic form, though there are caveats: we are working in an odd dimensional space so technically $\omega$ is not symplectic and $\theta$ for us is a vector potential, therefore the canonical one form/contact form correspond to a specific gauge choice of the vector potential. These are exactly the type of mismatches between the physics and the mathematics that we had to overcome before being able to connect everything. We believe the full mathematical account of these differences would clutter the exposition, so we hope the sketches we provide are sufficient.

\begin{quote}
There are some minor typos, but otherwise the language is fine and readable. In fact, the authors are gifted writers, and I appreciate that the paper is under 75 lines!
\end{quote}
We have addressed a handful of typos we found. We thank the referee for the kind words.

@@

Some of these ideas have been explored in our papers "Hamiltonian mechanics is conservation of information entropy" and "Hamiltonian privilege". We haven't cited them as we didn't want to have too many self-citations. We are also working on a new edition of our "Assumptions of Physics" book, which we have now cited in the acknowledgments, that will contain the full account for classical systems.

@@

\clearpage
\bibliographystyle{apalike}
\bibliography{bibliography}
\end{document}

>>>>

\begin{quote}
 Some of the standard notation and terminology of differential geometry are not present in the paper. For example, the discussion in the supplementary document from line 115 to 124 could be referred to as the volume preserving diffeomorphism of flow maps that are presented in Kambe, Tsutomu Jixin. Geometrical Theory Of Dynamical Systems And Fluid Flows (Revised Edition). Vol. 23. World Scientific Publishing Company, 2009. Also, for standard notation of differential geometry, please see Kreyszig, Erwin. Differential geometry. Courier Corporation, 2013. 
\end{quote}
As we stated in the Supplementary Information of the original version, the use of standard notation is avoided on purpose because, in our experience, it does not map well to physical intuition. We added an additional comment to that effect in the main body of the paper on line 26, as we realized this was not made sufficiently clear. In response to the other reviewer, we have also clarified in the paper the connection between the mathematical objects $\vec{S}$, $\omega$ and $\theta$ in our paper and in the standard mathematical literature, which further highlights the mismatch between standard mathematical definitions and the physically motivated ones. For example, $\theta$ would correspond to the canonical one-form in the context of symplectic geometry and the contact form in the context of contact manifolds. Both of these objects are uniquely defined mathematically. However, given that the physical object we represent with $\theta$ is a vector potential, it is not uniquely defined as it depends on the arbitrary gauge choice. Therefore the ``mathematical'' $\theta$s represent the vector potential with a specific gauge choice.

@@

\begin{quote}
 General comments and questions

 Is the study limited to divergence free vector fields only? 
\end{quote}
Yes. The study is limited to all cases where the principle of stationary action holds (i.e. yields unique solutions) which, by virtue of the Liouville theorem, corresponds to divergence free displacement vector fields.

@@

\begin{quote}
 Why call the Principle of Least Action the Action Principle? 
\end{quote}
We use the term in the title to avoid making it exceedingly long. It is not an uncommon abbreviation. For example, the book ``The action principle and partial differential equations'' by D. Christodoulou.  We generally use ``principle of stationary action'' throughout the text of the paper.

@@

 In pp. 11 line 140 in the supplementary document, “Putting …” should be “Summing or Grouping …”.
\end{quote}
We thank the reviewer for taking the time to point out the typos, which have been fixed. We also have reworded the sentence beginning ``Putting ...''.

@@

\begin{quote}
	Now, I am curious, and would like the authors to clarify what mathematical objects are new.  The displacement field S was new to me. Is the potential $\theta$, also new?  And the counting form, lower case $\omega$, is this the symplectic 2-form familiar from Hamiltonian geometry?  If so, please state. And if not, perhaps a different symbol would be better.
\end{quote}
We added remarks when the symbols are introduced in the main text, at lines 41, 52 and 98. We also added more explanation in the Supplementary Information at lines 186, 204 and 380. The displacement field S is typically referred to as ``the vector field'' in dynamical system theory. $\theta$ roughly corresponds to the canonical one-form/contact form and $\omega$ to the symplectic form, though there are caveats: we are working in an odd-dimensional space so technically $\omega$ is not symplectic and $\theta$ for us is a vector potential, therefore the canonical one-form/contact form correspond to a specific gauge choice of the vector potential. These are exactly the type of mismatches between the physics and the mathematics that we had to overcome before being able to connect everything. We believe the full mathematical account of these differences would clutter the exposition, so we hope the sketches we provide are sufficient.

\begin{quote}
There are some minor typos, but otherwise the language is fine and readable. In fact, the authors are gifted writers, and I appreciate that the paper is under 75 lines!
\end{quote}
We have addressed the handful of typos we found. We thank the referee for the kind words.

@@

Some of these ideas have been explored in our papers ``Hamiltonian mechanics is conservation of information entropy'' (Studies in History and Philosophy of Modern Physics 71:60, 2020) and ``Hamiltonian privilege'' (Erkenntnis 2023). We haven't cited them as we didn't want to have too many self-citations. We are also working on a new edition of our \emph{Assumptions of Physics} book (Michigan Publishing 2021), which we have now cited in the acknowledgments, that will contain the full account for classical systems.

@@

%\clearpage
%\bibliographystyle{apalike}
%\bibliography{bibliography}
\end{document}
